\sysfull

### 2017-11-17 paper/paging.tex
paging: numUncomitted -> numShadowed rename bug
sec 5 - typos

As they are
accessed by a task, pages are swapped from their home location
in the direct\hyp{}mapped non-volatile memory, called
\texttt{\underline{store}} (underline denotes allocation in non-volatile
memory) into the fully\hyp{}associative \texttt{working} page buffer in volatile memory. The
\texttt{working} buffer is allocated to take up the volatile memory that
remains after space for stack is allocated.
%
\sys uses two-phase commit to write dirty pages from the \texttt{working} buffer
back to their home location in \texttt{\underline{store}} atomically,
ensuring data consistency even with power failures.
%
When any dirty page is swapped out from the \texttt{working} buffer due to
eviction or task transition from last coalesced task, the first phase writes
the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}}
buffer, and the second phase writes it from
\texttt{\underline{shadow}} to \texttt{\underline{store}}.
%
Unline prior work~\cite{chain,alpaca} that moves scattered variables one at a
time between volatile and non-volatile memory, \sys's page-based
design moves contiguous blocks of data.
%
\sys takes full advantage of hardware\hyp{}accelerated block copies using
Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5x
fewer cycles on MSP430FRxxxx.

@@

\begin{algorithm}
	\caption{\texttt{PageFault}(tag $t$)}
	\label{algo:pagefault}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
        \State \LeftComment{$i_\text{victim}$: volatile index to next free slot or page to be evicted, set to 0 at boot}
        \State \LeftComment{\texttt{\underline{numShadowed}}: non-volatile, reset to 0 on boot unless in phase 2 commit}
        \State $t_\text{victim} \gets \Call{Tag}{\texttt{working}[i_\text{victim}]}$
		\If {\Call{IsDirty}{working[$i_\text{victim}$]} }	\Comment{Are we evicting a dirty page?}
		\State $\texttt{\underline{shadow}}[t_\text{victim}] \xleftarrow{\text{DMA}} \texttt{working}[i_\text{victim}]$
            \Comment{Swap out the page}
        \If {$t_\text{victim} \not\in \texttt{\underline{shadowList}}$} \Comment{Keep track as list for quick traversal}
            \State $\texttt{\underline{shadowList}}[\texttt{numUncommited++}] \gets t_\text{victim}$
        \EndIf
		\EndIf
		\If {\Call{IsValid}{\underline{shadow}[$t$]}} \Comment{Had the page been swapped out earlier?}
		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
            \Comment{Get page from commit buffer}
		\Else
		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{store}}[t]$
            \Comment{Get page from backing store}
		\EndIf 
        \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBuf}|$ \Comment{FIFO eviction policy}
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[t]
	\caption{Two-phase commit}
	\label{algo:commit}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
        \Procedure{CommitPhase1}{} \Comment{Invoked from \texttt{next\_task} on last coalesced task}
            \For {$i \in 0..|\texttt{working}|-1$}
                \State $t \gets \Call{Tag}{\texttt{working}[i]}$
                \State $\texttt{\underline{shadow}}[\Call{Tag}{\texttt{working}[i]}] \xleftarrow{\text{DMA}} \texttt{working}[i]$
                \State $\texttt{\underline{shadowList}}[\texttt{\underline{numShadowed}}] \gets t$
                \State $\texttt{\underline{numUncommited}} \gets \texttt{\underline{numUncommited}} + 1$
            \EndFor
            \State \Call{CommitPhase2}{}
        \EndProcedure
        \Procedure{CommitPhase2}{}
            \State $\texttt{\underline{committing}} \gets \textbf{true}$
            \While{\texttt{\underline{numUncommited}} > 0}
                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{numUncommited}} - 1]$
                \State $\texttt{\underline{store}}[t] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
                \State $\texttt{\underline{shadow}}[t] \gets \textbf{nil}$ \Comment{Clear entry for page fault handler}
                \State $\texttt{\underline{numUncommited}} \gets \texttt{\underline{numUncommited}} - 1 $
            \EndWhile
            \State $\texttt{\underline{committing}} \gets \textbf{false}$
        \EndProcedure
        \Procedure{OnBoot}{} \Comment{Invoked on every boot, before transitioning to a task}
            \If { \texttt{\underline{committing}} } \Call{CommitPhase2}{}
            \EndIf 
            \State $\texttt{\underline{numShadowed}} \gets 0$, $i_\text{victim} \gets 0$
        \EndProcedure
	\end{algorithmic}
\end{algorithm}

>>>>

As they are
accessed by a task, pages are swapped from their home location
in the direct\hyp{}mapped non-volatile memory, called
\texttt{\underline{store}} (underline denotes allocation in non-volatile
memory) into the fully\hyp{}associative \texttt{working} page buffer in volatile memory. The
\texttt{working} buffer is allocated to take up the volatile memory that
remains after space for stack is allocated.
%
\sys uses two-phase commit to write dirty pages from the \texttt{working} buffer
back to their home location in \texttt{\underline{store}} atomically,
ensuring data consistency even with power failures.
%
When any dirty page is swapped out from the \texttt{working} buffer due to
eviction or task transition from last coalesced task, the first phase writes
the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}}
buffer, and the second phase writes it from
\texttt{\underline{shadow}} to \texttt{\underline{store}}.
%
Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a
time between volatile and non-volatile memory, \sys's page-based
design moves contiguous blocks of data.
%
\sys takes full advantage of hardware\hyp{}accelerated block copies using
Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5x
fewer cycles on MSP430FRxxxx.

@@

\begin{algorithm}
	\caption{\texttt{PageFault}(tag $t$)}
	\label{algo:pagefault}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
        \State \LeftComment{$i_\text{victim}$: volatile index to next free slot or page to be evicted, set to 0 at boot}
        \State \LeftComment{\texttt{\underline{numShadowed}}: non-volatile, reset to 0 on boot unless in phase 2 commit}
        \State $t_\text{victim} \gets \Call{Tag}{\texttt{working}[i_\text{victim}]}$
		\If {\Call{IsDirty}{working[$i_\text{victim}$]} }	\Comment{Are we evicting a dirty page?}
		\State $\texttt{\underline{shadow}}[t_\text{victim}] \xleftarrow{\text{DMA}} \texttt{working}[i_\text{victim}]$
            \Comment{Swap out the page}
        \If {$t_\text{victim} \not\in \texttt{\underline{shadowList}}$} \Comment{Keep track as list for quick traversal}
            \State $\texttt{\underline{shadowList}}[\texttt{numShadowed++}] \gets t_\text{victim}$
        \EndIf
		\EndIf
		\If {\Call{IsValid}{\underline{shadow}[$t$]}} \Comment{Had the page been swapped out earlier?}
		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
            \Comment{Get page from commit buffer}
		\Else
		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{store}}[t]$
            \Comment{Get page from backing store}
		\EndIf 
        \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBuf}|$ \Comment{FIFO eviction policy}
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[t]
	\caption{Two-phase commit}
	\label{algo:commit}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
        \Procedure{CommitPhase1}{} \Comment{Invoked from \texttt{next\_task} on last coalesced task}
            \For {$i \in 0..|\texttt{working}|-1$}
                \State $t \gets \Call{Tag}{\texttt{working}[i]}$
                \State $\texttt{\underline{shadow}}[\Call{Tag}{\texttt{working}[i]}] \xleftarrow{\text{DMA}} \texttt{working}[i]$
                \State $\texttt{\underline{shadowList}}[\texttt{\underline{numShadowed}}] \gets t$
                \State $\texttt{\underline{numShadowed}} \gets \texttt{\underline{numShadowed}} + 1$
            \EndFor
            \State \Call{CommitPhase2}{}
        \EndProcedure
        \Procedure{CommitPhase2}{}
            \State $\texttt{\underline{committing}} \gets \textbf{true}$
            \While{\texttt{\underline{numShadowed}} > 0}
                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{numShadowed}} - 1]$
                \State $\texttt{\underline{store}}[t] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
                \State $\texttt{\underline{shadow}}[t] \gets \textbf{nil}$ \Comment{Clear entry for page fault handler}
                \State $\texttt{\underline{numShadowed}} \gets \texttt{\underline{numShadowed}} - 1 $
            \EndWhile
            \State $\texttt{\underline{committing}} \gets \textbf{false}$
        \EndProcedure
        \Procedure{OnBoot}{} \Comment{Invoked on every boot, before transitioning to a task}
            \If { \texttt{\underline{committing}} } \Call{CommitPhase2}{}
            \EndIf 
            \State $\texttt{\underline{numShadowed}} \gets 0$, $i_\text{victim} \gets 0$
        \EndProcedure
	\end{algorithmic}
\end{algorithm}
